Inferring the distribution and rate of Earth-like exoplanets \\ from noisy individual-planet inferences

### 2014-05-13 document/ms.tex
updated figures without NaN errorbars
notes about priors

The constraints in \eq{crazypost} can always be---and often are---propagated
as a list of $N$ samples $\{\entry_k^{(n)}\}$ from the posterior
\begin{eqnarray}
\entry_k^{(n)} &\sim& p(\{\entry_k\}\,|\,\{\data_k\},\,\interim) \quad.
\end{eqnarray}
We can use these samples to approximately compute the integral in
\eq{simplemarglike} as
\begin{eqnarray}\eqlabel{importance}
\frac{p(\{\data_k\}\,|\,\ratepars)}{p(\{\data_k\}\,|\,\interim)} &\approx&
    \frac{1}{N} \sum_{n=1}^N
    \frac{p(\{\entry_k^{(n)}\}\,|\,\ratepars)}
         {p(\{\entry_k^{(n)}\}\,|\,\interim)} \quad.
\end{eqnarray}
This is very efficient to compute as long as an evaluation of
$p(\{\entry_k^{(n)}\}\,|\,\ratepars)$ is not very expensive.
That being said, this could be a high variance estimator depending on the
initial choice of $p(\{\entry_k^{(n)}\}\,|\,\interim)$ but in our case, we'll
show that not many samples are needed to get an accurate estimate.

@@

@@

\paragraph{Occurence rate parameterization \& priors}
For the remainder of this \paper, we will also model the occurence rate as a
histogram in logarithmic bins of period $\ln\period$ and radius $\ln\radius$.

@@

As well as inferring the ocurrence distribution of exoplanets, this dataset
can also be used to constrain the rate of Earth analogs.
Even though no true analog has been observed, we can make some assumptions
about the distribution (as described below) and extrapolate to the location of
the Earth.
Explicitly, we constrain the rate of exoplanets \emph{evaluated at the
location of Earth}
\begin{eqnarray}
\gammaearth &=& \rate (\ln\period_\oplus,\,\ln\radius_\oplus) \\
&=&
\left.\frac{\dd N}{\dd\ln\period\,\dd\ln\radius}\right|
_{\radius=\radius_\oplus,\,\period=\period_\oplus}\quad.
\end{eqnarray}

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=0.7\textwidth]{figures/smooth/rate.pdf}
\end{center}
\caption{%
The extrapolated rate of Earth analogs \gammaearth\ as inferred by the
different techniques.
Applying the method used by \citet{petigura} gives a constraint indicated by
the vertical black line with error bars shown as dashed lines.
The histogram is the MCMC estimate of our posterior constraint on this rate
and the true value is indicated as the thick gray vertical line.
\figlabel{smooth-rate}}
\end{figure}

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=0.7\textwidth]{figures/simulation/rate.pdf}
\end{center}
\caption{%
The same as \fig{smooth-rate} for \modelb.
\dfmtodo{UPDATE THIS FIGURE TO FINAL VERSION}
\figlabel{simulation-rate}}
\end{figure}

@@

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=0.7\textwidth]{figures/results/rate.pdf}
\end{center}
\caption{%
The same as \fig{smooth-rate} but applied to the catalog from \citet{petigura}.
\figlabel{real-rate}}
\end{figure}

>>>>

The constraints in \eq{crazypost} can always be---and often are---propagated
as a list of $N$ samples $\{\entry_k^{(n)}\}$ from the posterior
\begin{eqnarray}
\{\entry_k\}^{(n)} &\sim& p(\{\entry_k\}\,|\,\{\data_k\},\,\interim) \quad.
\end{eqnarray}
We can use these samples to approximately compute the integral in
\eq{simplemarglike} as
\begin{eqnarray}\eqlabel{importance}
\frac{p(\{\data_k\}\,|\,\ratepars)}{p(\{\data_k\}\,|\,\interim)} &\approx&
    \frac{1}{N} \sum_{n=1}^N
    \frac{p(\{\entry_k^{(n)}\}\,|\,\ratepars)}
         {p(\{\entry_k^{(n)}\}\,|\,\interim)} \quad.
\end{eqnarray}
This is very efficient to compute as long as an evaluation of
$p(\{\entry_k^{(n)}\}\,|\,\ratepars)$ is not very expensive.
That being said, this could be a high variance estimator depending on the
initial choice of $p(\{\entry_k^{(n)}\}\,|\,\interim)$ but in our case, we'll
show that not many samples are needed to get an accurate estimate.

@@

Now, we can substitute \eq{poisson-like} into \eq{simplemarglike} and apply
the importance sampling approximation to derive the following expression for
the marginalized likelihood
\begin{eqnarray}\eqlabel{money}
\frac{p(\{\data_k\}\,|\,\ratepars)}{p(\{\data_k\}\,|\,\interim)} &\approx&
    \exp\left(-\int \obs{\rate}_\ratepars (\entry) \dd\entry\right) \,
    \prod_{k=1}^K
    \frac{1}{N_k} \sum_{n=1}^{N_k}
    \frac{\obs{\rate}_\ratepars (\entry_k^{(n)})}
         {p(\entry_k^{(n)}\,|\,\interim)} \quad.
\end{eqnarray}
In this equation, we're making the further assumption that the catalog treated
the objects independently.
In other words, we have per-object posterior samples
\begin{eqnarray}
\entry_k^{(n)} &\sim& p(\entry_k\,|\,\data_k,\,\interim) \quad.
\end{eqnarray}
\Eq{money} is the \emph{money equation} for our method.
It lets us efficiently compute the \emph{marginalized likelihood of the entire
set of light curves for a particular occurence rate function}.

@@

\paragraph{Occurence rate parameterization \& priors}
For the remainder of this \paper, we also model the occurence rate as a
two-dimensional histogram with fixed logarithmic bins in period and radius.
When we include observational uncertainties---using \eq{money}---the maximum
likelihood result is no longer analytic.
Therefore, if we want to compute the ``best-fit'' occurence rate, we can use a
standard non-linear optimization algorithm.

@@

As well as inferring the ocurrence distribution of exoplanets, this dataset
can also be used to constrain the rate of Earth analogs.
Even though no true analog has been observed, we can make some assumptions
about the distribution (as described below) and extrapolate to the location of
the Earth.
Explicitly, we constrain the rate of exoplanets \emph{evaluated at the
location of Earth}
\begin{eqnarray}
\gammaearth &=& \rate (\ln\period_\oplus,\,\ln\radius_\oplus) \\
&=&
\left.\frac{\dd N}{\dd\ln\period\,\dd\ln\radius}\right|
_{\radius=\radius_\oplus,\,\period=\period_\oplus}\quad.
\end{eqnarray}
This is also the quantity that \citet{petigura} estimated in different units.

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=0.6\textwidth]{figures/smooth/rate.pdf}
\end{center}
\caption{%
The extrapolated rate of Earth analogs \gammaearth\ as inferred by the
different techniques.
Applying the method used by \citet{petigura} gives a constraint indicated by
the vertical black line with error bars shown as dashed lines.
The histogram is the MCMC estimate of our posterior constraint on this rate
and the true value is indicated as the thick gray vertical line.
\figlabel{smooth-rate}}
\end{figure}

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=0.6\textwidth]{figures/simulation/rate.pdf}
\end{center}
\caption{%
The same as \fig{smooth-rate} for \modelb.
\dfmtodo{UPDATE THIS FIGURE TO FINAL VERSION}
\figlabel{simulation-rate}}
\end{figure}

@@

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=0.6\textwidth]{figures/results/rate.pdf}
\end{center}
\caption{%
The same as \fig{smooth-rate} but applied to the catalog from \citet{petigura}.
\figlabel{real-rate}}
\end{figure}
